---
title: "06_update"
author: "randy"
date: '2022-04-19'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


```{r include=FALSE}
library(tidyverse)
library(magrittr)

library(MASS)
library(here)

library(brokenstick)
library(gamlss)
library(lme4)

library(data.table)
library(gtsummary)
library(flextable)
library(splines)


library(ggridges)
```


1. Redo table 1 with final sample  

```{r include=FALSE}
table1 <- here::here("data", "epic_table1.csv") %>%
  read.csv(row.names = 1) %>%
  unique() %>%
  dplyr::select(-id) %>%
  mutate(
    ethnic = case_when(ethnic == 1 ~ "Hispanic",
                     ethnic == 2 ~ "Non-Hispanic"),
    race = case_when(race == 1 ~ "White",
                     race != 1 ~ "Other"),
    sex = case_when(sex == "F" ~ "Female",
                    sex == "M" ~ "Male"),
    age_diff = age_max - age_min) %>% 
  dplyr::select(group,
         Genotype = genotype,
         Gender = sex,
         Race = race,
         Ethnicity = ethnic,
         "Visit number" = visitn,
         "Age at registration, years" = age_min, 
         "Age at end of follow-up, years" = age_max, 
         "Follow-up, years" = age_diff,
         "Height at registration, cm" = h_min,
         # "Height mean" = h_mean, 
         # "Weight mean" = w_mean,
         "Weight at registration, kg" = w_min) %>% 
  ## select all the variables for table1
  tbl_summary(by = group) %>%
  ## just display all the variables in one column
  modify_header(label = "**Variable**") %>%
  # update the column header
  bold_labels() %>%
  italicize_labels() %>%
  as_flex_table() %>%
  flextable::bold(part = "header") %>% 
  ## auto adjust the column widths 
  flextable::autofit()
```


```{r}
table1

## save pptx -----------------------------------------------
## flextable can be saved directly to powerpoints
flextable::save_as_pptx(
  table1, 
  path = "figure/01_table1.pptx")
```


2. Refine spline model for training dataset. 

Try spline model with 3 inner knots set at percentiles (quartiles) too 
and compare to the splines you had tried before. 

```{r}
train <- here::here("data", "epic_clean_randy.csv") %>%
  read.csv(row.names = 1) %>% 
  filter(group == "training")

test <- here::here("data", "epic_clean_randy.csv") %>%
  read.csv(row.names = 1) %>%
  filter(group == "testing")

baseline <- train %>% 
  group_by(id) %>% 
  arrange(time) %>% 
  slice(1L) %>%
  rename(ht_base = ht,
         time_base = time) %>%
  dplyr::select(id, ht_base, time_base)
```


```{r}
# lmeControl
ctrl <- lmeControl(opt='optim')

## the original mod2_2_4 in file "04_model_fit.Rmd"
fit2_0 <- lme(ht ~ bs(time, knots = c(10, 15), 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit2_1 <- lme(ht ~ bs(time, knots = c(10, 12), 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit2_2 <- lme(ht ~ bs(time, df = 5, 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)

fit3_0 <- lme(ht ~ bs(time, df = 6, 
                      degree = 3,
                      intercept = FALSE) * sex +  genotype - 1,
              control = ctrl,
              random = ~ 1| id,
              data = train)
```


```{r}
get_tag <- function(fit, data1 = train) {
  tidy <- broom.mixed::tidy(fit)
  glance <- broom.mixed::glance(fit)
  augment <- broom.mixed::augment(fit) %>%
    dplyr::select(time, .fitted, sex) %>%
    unique()
  
  plot1 <- ggplot() +
        geom_smooth(data = data1, 
                    aes(x = time, y = ht, 
                        group = sex, color = sex)) +
        geom_smooth(data = augment, 
                    aes(x = time, y = .fitted, 
                        group = sex),
                    linetype = 2) +
        xlab("Age") +
        theme_bw()


  return(list(tidy = tidy, 
              glance = glance, 
              augment = augment,
              plot = plot1))
}
```

```{r}
tag2_0 <- get_tag(fit = fit2_0)
tag2_1 <- get_tag(fit = fit2_1)
tag2_2 <- get_tag(fit = fit2_2)
tag3_0 <- get_tag(fit = fit3_0)

glance_all <- map_dfr(list(tag2_0, tag2_1,
                    tag2_2, tag3_0),
                  "glance") %>%
  mutate(model = c("fit2_0", "fit2_1", "fit2_2", "fit3_0")) %>%
  dplyr::select(model, everything())

glance_all
```

3. plot the predicted/expected means. 

```{r echo=FALSE, fig.height=9, fig.width=10, message=FALSE, warning=FALSE}
# plot_all <- map(list(tag2_0, tag2_1,
#                     tag2_2, tag3_0),
#                   "plot")
# unlist(plot_all)
library("gridExtra")
grid.arrange(tag2_0$plot, tag2_1$plot,
             tag2_2$plot, tag3_0$plot,
             ncol = 2, nrow = 2)
```

4. For brokenstick model try time for a anchoring of 10 and 12  

```{r}
bks1 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    subjid = train$id,
                    knots = c(10, 15))
est1_knots <- predict(bks1, 
                     x = "knots", 
                     group = train$id)
est1_all <- predict(bks1, 
                   group = train$id)


bks2 <- brokenstick(formula = ht ~ time | id,
                    data = train,
                    subjid = train$id,
                    knots = c(10, 12))
est2_knots <- predict(bks2, 
                     x = "knots", 
                     group = train$id)
est2_all <- predict(bks2, 
                   group = train$id)
```


```{r fig.height=5, fig.width=5}
plot_bks1 <- ggplot() +
  geom_smooth(data = train, 
              aes(x = time, y = ht, 
                  group = sex),
              alpha = 0.3,
              size = 1,
              color = "darkgreen") +
  geom_smooth(data = est1_all, 
              aes(x = time, y = .pred, 
                  group = sex),
              color = "brown",
              linetype = 2,
              size = 0.5) +
  geom_smooth(data = est2_all, 
                    aes(x = time, y = .pred, 
                        group = sex),
              color = "magenta",
              linetype = 4,
              size = 0.5) +
  xlab("Age") +
  theme_bw()

plot_bks1
```


5. Make progress, as able with the latex file of the paper. 
Start drafting a description of the dataset using Moss and Rice’s papers: 
one paragraph to describe the study and a second to include descriptive statistics.   
 


